dispRity manual

### 2017-01-03 doc/dispRity-manual.Rmd
further man updates

## What is in the object?
names(median_centroids)

@@

```{r}
## This object contains three elements
names(crown_stem)
## With "matrix" being the original matrix (the ordinated space)
str(crown_stem$matrix)
## And "series" the elements from this ordinated space into the two sub-spaces
str(crown_stem$series)
## Note that the elements here are indicated as numeric values referring to the row numbers in the
## original matrix
```

@@

This function intakes a `dispRity` object plus various options namely the `quantile` values for the confidence intervals levels; the `cent.tend` for the 
central tendency to use for summarising the results and two *visual* options which are whether to recall the `dispRity` options and how much digits are wanted in the results.
By default the 50 and 95 quantiles and the median are calculated but these options can be modified as follows:

<!-- STOPPED HERE -->

```{r}
## Summarising the disparity using the 88 quantile and the standard deviation as the central
## tendency. 
summary(disp_time_binsBS, quantile = 88, cent.tend = sd)

## If one happens to forgot what was calculated in the summarised object,
## it is possible to recall the different steps details using recall:
summary(disp_time_binsBS, recall = TRUE)

## Note that the information with each different number of elements is
## displayed for rarefied data:
tail(summary(disp_crown_stemBS))
## The observed disparity is NA for rarefied data.
```

@@

An alternative way to summarise the data is to plot the results using the S3 method `plot.dispRity`.
This function intakes the same options as `summary.dispRity` along side with various graphical options described in the function manual (see `?plot.dispRity`).

Two main categories of plots can be displayed: continuous plots with the argument `type = continuous` (see below) and discrete plots (default is `type = box`.
Note that this argument can be left missing. In this case, the algorithm will automatically detect the type of series from the `dispRity` object.

The discrete plot arguments can be either:

-   `type = box` for plotting boxplots.

-   `type = polygon` for plotting boxes that represent the different levels of confidence intervals (with a dot for the central tendency).

-   `type = lines` which is similar to `type = polygon` but plots lines rather than boxes for the confidence intervals levels.

It is also possible to display the number of elements per series (has a horizontal doted line) using the option `elements = TRUE`.
Additionally, when the data is rarefied (see section \[rare\]), one can also indicate which level of rarefaction to display (i.e. display the results for a certain number of elements) by using the `rarefaction` argument.

```{r, fig.width=6, fig.height=6}
## Graphical parameters
op <- par(mfrow = c(2,2), bty = "n")

## Plotting the disparity for both groups
plot(disp_crown_stemBS, type = "box")

## Same plot but with the number of elements on the right and using the
## "polygon" plot type.
plot(disp_crown_stemBS, type = "polygon", elements = TRUE)
## Rarefied version of the same plot (20 elements per series)
plot(disp_crown_stemBS, type = "polygon", rarefaction = 20)
```

It is also possible to easily look at the behaviour of the disparity metric with the rarefied data (i.e. how does the disparity score change when the number of elements goes down) by using the `rarefaction = plot` argument:

```{r, fig.width=6, fig.height=6}
## Graphical options
par(bty = "n")
## The rarefaction curves
plot(disp_crown_stemBS, elements = TRUE, rarefaction = "plot")

@@

Some of the core functions from the package have a `parallel` option.

>>>>

## What does the object contain?
names(median_centroids)

@@

```{r}
## This object contains three elements
names(crown_stem)
## With "matrix" being the original matrix (the ordinated space)
str(crown_stem$matrix)
## And "series" the elements from this ordinated space into the two sub-spaces
str(crown_stem$series)
```

@@

This function intakes a `dispRity` object plus various options namely the `quantile` values for the confidence intervals levels (by default, the 50 and 95 quantiles are calculated); the `cent.tend` for the  central tendency to use for summarising the results (default is `median`); the `rounding` option corresponding to the number of decimals to print (default is `2`) and the `recall` option for printing the call of the `dispRity` object as well (default is `FALSE`)
These options can easily be changed from the defaults as follows:

```{r}
## Summarising the results using the 88 quantile and the standard deviation as the central tendency. 
summary(disp_time_binsBS, quantile = 88, cent.tend = sd)

## Printing the details of the dispRity object and rounding the values to the 5th decimal
summary(disp_time_binsBS, recall = TRUE, rounding = 5)
```

> Information about the number of elements per series as well as the observed (i.e. non-bootstrapped) disparity is also calculated.
This is specifically handy when rarefying the data for example

```{r}
head(summary(disp_crown_stemBS))
```

@@

An alternative way to display the calculated disparity is to plot the results using the S3 method `plot.dispRity`.
This function intakes the same options as `summary.dispRity` along side with various graphical options described in the function manual (see `?plot.dispRity`).

The plots can be of four different types: `continuous` for displaying continuous disparity curves and `box`, `lines` and `polygons` to display discrete disparity results in respectively a boxplot, confidence interval lines and confidence interval polygons.

> This argument can be left missing. In this case, the algorithm will automatically detect the type of series from the `dispRity` object.

It is also possible to display the number of elements per series (has a horizontal doted line) using the option `elements = TRUE`.
Additionally, when the data is rarefied, one can also indicate which level of rarefaction to display (i.e. display the results for a certain number of elements) by using the `rarefaction` argument.

```{r, fig.width=6, fig.height=6}
## Graphical parameters
op <- par(mfrow = c(2,2), bty = "n")
## Plotting continuous disparity results
plot(disp_time_binsBS, type = "continuous")
## Plotting discrete disparity results
plot(disp_crown_stemBS, type = "box")
## Plotting the disparity using lines for the rarefaction level of 20 elements only
plot(disp_crown_stemBS, type = "lines", rarefaction = 20)
## Plotting the disparity using polygons while also displaying the number of elements
plot(disp_crown_stemBS, type = "polygon", elements = TRUE)
```

> The `type` argument is not mandatory for the two first plots since both `continuous` and `box` are the default options for respectively continuous and discrete data.

It is also possible to easily look at the behaviour of the disparity metric with the rarefied data (i.e. how does the disparity score change when the number of elements goes down) by using the `rarefaction = TRUE` argument:

<!-- 
```{r, fig.width=6, fig.height=6}
## Graphical options
par(bty = "n")
## The rarefaction curves
plot(disp_crown_stemBS, elements = TRUE, rarefaction = TRUE)

@@

Some of the core functions from the package have a `parallel` option.
 -->
